ODGI: scalable tools for pangenome graphs

### 2021-06-02 manuscript/main.tex
edit abstract

    \abstract{ \textbf{Motivation:} Pangenomes graphs address some of the
      shortcomings of mainline genomics, in particular
      with regard to reference genome bias and identification of complex
      structural variants.
        %
        \\ \textbf{Results:} Here we present \odgi\, a novel range of
        scalable tools that implements new algorithms and has an efficient in-memory
        representation of DNA graphs. \odgi\ can handle hundreds of
        large genomes in memory with fast
        parallel execution.  \odgi\ includes
        tools for detecting complex regions, extracting \textit{loci},
        exploratory analysis, sorting, removing artifacts, and graph
        manipulation and visualisation. \odgi\ also provides an extensible API that allows
        for writing new tool.
        %
        \\
        \textbf{Availability:} \odgi\ is software written in the \Cpp\ programming language and
        published under the permissive MIT license. Source code and documentation can be downloaded from
        \href{https://github.com/pangenome/odgi}{https://github.com/pangenome/odgi}.\\
        %
        \textbf{Contact:}
        % \href{andreaguarracino@outlook.com}{andreaguarracino@outlook.com},
        egarris5@uthcs.edu
        \textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online. \FIXME
        }
    \maketitle

@@

    Thanks to advances in sequencing technology, new full-length
    genome assemblies are being produced at a high rate, including
    efforts from the Human Pangenome Reference Consortium (HPRC) and
    telomere-to-telemere projects, e.g.~\citep{Miga:2020}.  In
    principle, a pangenome can model a full set of genomic elements of
    individuals in a given species or clade. Pangenomics contrasts
    with reference genome based approaches which relate sequences to a
    single consensus model of the genome~\citep{Eizenga:2020}. A
    pangenome can elegantly be represented by a graph data structure
    incorporating sequences as nodes and their connections as
    edges. These nodes are shared for homologs, paralogs and
    orthologs. So-called paths represent reads, contigs, haplotypes, individual chromosomes or part thereof, that run through the
    nodes.   A bi-directed graph can even contain both strands of DNA
    as paths. A graph-based approach is particularly relevant for
    exploring complex variation and highly repetitive regions.

>>>>

    \abstract{ \textbf{Motivation:} Pangenomes graphs address some of the
      shortcomings of mainline genomics, in particular
      with regard to reference genome bias and identification of complex
      structural variants.
        %
        \\ \textbf{Results:} Here we present \odgi\, a novel range of
        scalable tools that implements new algorithms and has an efficient in-memory
        representation of DNA graphs. \odgi\ can handle hundreds of
        large genomes in memory with fast parallel execution. \odgi\ includes
        tools for detecting complex regions, extracting \textit{loci}, removing artifacts,
        exploratory analysis, and graph sorting, manipulation and visualisation.
        \odgi\ also provides an extensible API that allows for writing new tool.
        %
        \\
        \textbf{Availability:} \odgi\ is software written in the \Cpp\ programming language and
        published under the permissive MIT license. Source code and documentation can be downloaded from
        \href{https://github.com/pangenome/odgi}{https://github.com/pangenome/odgi}.\\
        %
        \textbf{Contact:}
        % \href{andreaguarracino@outlook.com}{andreaguarracino@outlook.com},
        egarris5@uthcs.edu
        \textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online. \FIXME
        }
    \maketitle

@@

    Thanks to advances in sequencing technology, new full-length
    genome assemblies are produced at a high rate, including
    efforts from the Human Pangenome Reference Consortium (HPRC) and
    telomere-to-telemere projects, e.g.~\citep{Miga:2020}. In
    principle, a pangenome can model a full set of genomic elements of
    individuals in a given species or clade. Pangenomics contrasts
    with reference genome based approaches which relate sequences to a
    single consensus model of the genome~\citep{Eizenga:2020}. A
    pangenome can elegantly be represented by a graph data structure
    incorporating sequences as nodes and their connections as
    edges. These nodes are shared for homologs, paralogs and
    orthologs. So-called paths represent reads, contigs, haplotypes, individual chromosomes or part thereof, that run through the
    nodes.   A bi-directed graph can even contain both strands of DNA
    as paths. A graph-based approach is particularly relevant for
    exploring complex variation and highly repetitive regions.
